Bibliography file created

\begin{thebibliography}{99}

\bibitem{mustafa2023dhaka}
Kallol Mustafa, ``Why exactly is Dhaka the slowest city in the world?'' \textit{The Daily Star}, Oct. 8, 2023. [Online]. Available: https://www.thedailystar.net/opinion/views/news/why-exactly-dhaka-the-slowest-city-the-world-3436751

\bibitem{karim2022traffic}
T. J. Karim, ``Traffic Gridlock: Time Wasted in Dhaka,'' \textit{Dhaka Tribune}, Dec. 5, 2022.

\bibitem{wu2014emergency}
J. D. J. Wu, M. R. Bell, and M. T. Williams, ``The Effect of Traffic Congestion on Emergency Vehicle Response Times,'' \textit{Journal of Transportation Engineering}, vol. 140, no. 8, 2014.

\bibitem{webster1958traffic}
F. V. Webster, ``Traffic Signal Settings,'' \textit{Road Research Technical Paper}, no. 39, Road Research Laboratory, HMSO, London, 1958.

\bibitem{rahman2020traffic}
M. M. Rahman and M. Mohiuddin, ``Traffic Management in Dhaka City: A Critical Review,'' \textit{Journal of Urban Planning and Development}, vol. 146, no. 1, p. 04019029, 2020.

\bibitem{hunt1981scoot}
P. B. Hunt, D. I. Robertson, R. D. Bretherton, and R. I. Winton, ``SCOOT-a traffic responsive method of coordinating signals,'' \textit{TRRL Laboratory Report}, no. 1014, Transport and Road Research Laboratory, Crowthorne, 1981.

\bibitem{zhang2020intelligent}
D. Zhang, Y. Wang, and X. Liu, ``Intelligent Traffic Light Control Using Deep Reinforcement Learning with Object Detection,'' \textit{IEEE Transactions on Intelligent Transportation Systems}, vol. 20, no. 3, pp. 1-10, 2020.

\bibitem{redmon2016yolo}
J. Redmon, S. Divvala, R. Girshick, and A. Farhadi, ``You Only Look Once: Unified, real-time object detection,'' in \textit{Proceedings of the IEEE Conference on Computer Vision and Pattern Recognition (CVPR)}, pp. 779-788, 2016.

\bibitem{singh2021realtime}
K. R. K. Singh, P. S. K. Prasad, and M. S. P. K. Roy, ``Real-time Traffic Monitoring and Analysis Using YOLOv3 Model,'' \textit{Journal of Intelligent Transportation Systems}, vol. 25, no. 5, pp. 509-520, 2021.

\bibitem{li2017diffusion}
Y. Li, R. Yu, C. Shahabi, and Y. Liu, ``Diffusion convolutional recurrent neural network: Data-driven traffic forecasting,'' in \textit{Proceedings of the International Conference on Learning Representations (ICLR)}, 2017.

\bibitem{farooq2020priority}
M. Farooq, M. A. Habib, and M. Iqbal, ``Priority-based Traffic Management System for Emergency Vehicles in Urban Areas,'' \textit{International Journal of Computer Applications}, vol. 175, no. 20, pp. 45-50, 2020.

\bibitem{wong2020intelligent}
K. Wong, L. Chen, and J. Zhang, ``Intelligent Traffic Signal Control System for Emergency Vehicle Priority,'' \textit{IEEE Transactions on Vehicular Technology}, vol. 69, no. 8, pp. 8324-8335, 2020.

\bibitem{sharma2021smart}
N. Sharma, A. Garg, and P. Jain, ``Smart traffic light control system using Raspberry Pi and IoT,'' \textit{International Journal of Computer Science and Network Security}, vol. 21, no. 3, pp. 158-164, 2021.

\bibitem{ahmed2019urban}
K. Ahmed and M. S. Rahman, ``Urban Traffic Congestion in Dhaka City: An Empirical Study,'' \textit{International Journal of Traffic and Transportation Engineering}, vol. 8, no. 2, pp. 19-30, 2019.

\bibitem{islam2020design}
M. S. Islam, S. A. Chowdhury, and M. A. Hossain, ``Design and implementation of an intelligent traffic control system using Arduino and machine learning,'' \textit{International Journal of Intelligent Systems and Applications}, vol. 12, no. 4, pp. 42-50, 2020.

\bibitem{deccanherald2023ai}
Deccan Herald, ``AI-powered signals in Bengaluru reduce travel time by 33\%,'' Mar. 12, 2023. [Online]. Available: https://www.deccanherald.com/city/bengaluru/ai-powered-signals-in-bengaluru-reduce-travel-time-by-33-1199234.html

\bibitem{yao2020}
Y. J. Yao, R. W. Yang, and L. Liu, "Design of Intelligent Traffic Management System Based on Video Detection Technology," \textit{IEEE Access}, vol. 8, pp. 67278-67289, 2020.

\bibitem{zhang2020}
Zhang, D., Wang, Y., \& Liu, X., "Intelligent Traffic Light Control Using Deep Reinforcement Learning with Object Detection," \textit{IEEE Transactions on Intelligent Transportation Systems}, vol. 20, no. 3, pp. 1-10, 2020.

\bibitem{wu2014}
J. D. J. Wu, M. R. Bell, and M. T. Williams, "The Effect of Traffic Congestion on Emergency Vehicle Response Times," \textit{Journal of Transportation Engineering}, vol. 140, no. 8, 2014. [Online]. Available: \url{https://ascelibrary.org/doi/abs/10.1061/(ASCE)TE.1943-5436.0000687}

\bibitem{farooq2020}
Farooq, M., Habib, M. A., \& Iqbal, M., "Priority-based Traffic Management System for Emergency Vehicles in Urban Areas," \textit{International Journal of Computer Applications}, vol. 175, no. 20, pp. 45-50, 2020.

\bibitem{ahmed2019}
Ahmed, K., \& Rahman, M. S., "Urban Traffic Congestion in Dhaka City: An Empirical Study," \textit{International Journal of Traffic and Transportation Engineering}, vol. 8, no. 2, pp. 19-30, 2019.

\bibitem{hossain2017}
MM Hossain \& A Kroeger, "87 Transport, delay to care and patient experience in pre-clinical emergency systems in dhaka city, bangladesh: a mixed methods study," 2017. [Online]. Available: \url{https://shorturl.at/CpxTO}

\bibitem{deccan2023}
Deccan Herald, "AI-powered signals in Bengaluru reduce travel time by 33\%," Mar. 12, 2023. [Online]. Available: \url{https://shorturl.at/EEUsQ}

\bibitem{hossain2017}
MM Hossain \& A Kroeger, "87 Transport, delay to care and patient experience in pre-clinical emergency systems in dhaka city, bangladesh: a mixed methods study," 2017. [Online]. Available: \url{https://shorturl.at/CpxTO}

\bibitem{sharma2021}
N. Sharma, A. Garg, and P. Jain, "Smart traffic light control system using Raspberry Pi and IoT," \textit{International Journal of Computer Science and Network Security}, vol. 21, no. 3, pp. 158-164, 2021.

\bibitem{islam2020}
M. S. Islam, S. A. Chowdhury, and M. A. Hossain, "Design and implementation of an intelligent traffic control system using Arduino and machine learning," \textit{International Journal of Intelligent Systems and Applications}, vol. 12, no. 4, pp. 42-50, 2020.

\bibitem{huang2019}
C. Huang, Q. Liu, and D. Zhang, "YOLO-based traffic signal control system using deep reinforcement learning," \textit{Transportation Research Part C: Emerging Technologies}, vol. 105, pp. 159-174, 2019.

\bibitem{gupta2020}
A. A. Gupta and S. R. Singh, "Impact of traffic congestion on emergency services in urban areas," \textit{International Journal of Transportation Science and Technology}, vol. 10, no. 2, pp. 123-135, 2020.

\bibitem{jocher2022}
G. Jocher, A. Chaurasia, A. Stoken, J. Borovec, NanoCode012, Y. Kwon, TaoXie, J. Fang, imyhxy, K. Michael, L. Changyu, J. Nadar, J. Laughing, UnglvKitDe, tkianai, yxNONG, P. Skalski, A. Hogan, D. Strobel, M. Jain, M. Mammana, A. Xie, D. Bertonha, J. Fati, X. Yifu, J. Chaurasia, R. Xie, J. Quach, \& U. Rai, "ultralytics/yolov5: v6.2 - YOLOv5 Classification Models, Apple M1, Reproducibility, ClearML and Deci.ai integrations," Aug. 2022. [Online]. Available: \url{https://doi.org/10.5281/zenodo.3908559}

\bibitem{wang2023}
C. Wang, A. Bochkovskiy, and H. M. Liao, "YOLOv7: Trainable bag-of-freebies sets new state-of-the-art for real-time object detectors," \textit{Proceedings of the IEEE/CVF Conference on Computer Vision and Pattern Recognition}, pp. 7464-7475, 2023.

\bibitem{chen2021}
X. Chen, P. Wang, and Z. Zhang, "Adaptive traffic signal control with deep reinforcement learning: A survey," \textit{IEEE Transactions on Intelligent Transportation Systems}, vol. 22, no. 12, pp. 7670-7685, 2021.

\bibitem{liu2020}
Z. Liu, Y. Lin, Y. Cao, H. Hu, Y. Wei, Z. Zhang, S. Lin, and B. Guo, "Swin transformer: Hierarchical vision transformer using shifted windows," \textit{Proceedings of the IEEE/CVF international conference on computer vision}, pp. 10012-10022, 2021.

\bibitem{bochkovskiy2020}
A. Bochkovskiy, C. Y. Wang, and H. Y. M. Liao, "YOLOv4: Optimal speed and accuracy of object detection," \textit{arXiv preprint arXiv:2004.10934}, 2020.

\bibitem{tan2020}
M. Tan, R. Pang, and Q. V. Le, "EfficientDet: Scalable and efficient object detection," \textit{Proceedings of the IEEE/CVF conference on computer vision and pattern recognition}, pp. 10781-10790, 2020.

\bibitem{ren2017}
S. Ren, K. He, R. Girshick, and J. Sun, "Faster R-CNN: Towards real-time object detection with region proposal networks," \textit{IEEE transactions on pattern analysis and machine intelligence}, vol. 39, no. 6, pp. 1137-1149, 2017.

\bibitem{lin2017}
T. Y. Lin, P. Goyal, R. Girshick, K. He, and P. Dollár, "Focal loss for dense object detection," \textit{Proceedings of the IEEE international conference on computer vision}, pp. 2980-2988, 2017.

\bibitem{dosovitskiy2020}
A. Dosovitskiy, L. Beyer, A. Kolesnikov, D. Weissenborn, X. Zhai, T. Unterthiner, M. Dehghani, M. Minderer, G. Heigold, S. Gelly, J. Uszkoreit, and N. Houlsby, "An image is worth 16x16 words: Transformers for image recognition at scale," \textit{arXiv preprint arXiv:2010.11929}, 2020.

\bibitem{he2016}
K. He, X. Zhang, S. Ren, and J. Sun, "Deep residual learning for image recognition," \textit{Proceedings of the IEEE conference on computer vision and pattern recognition}, pp. 770-778, 2016.

\bibitem{simonyan2014}
K. Simonyan and A. Zisserman, "Very deep convolutional networks for large-scale image recognition," \textit{arXiv preprint arXiv:1409.1556}, 2014.

\bibitem{howard2017}
A. G. Howard, M. Zhu, B. Chen, D. Kalenichenko, W. Wang, T. Weyand, M. Andreetto, and H. Adam, "MobileNets: Efficient convolutional neural networks for mobile vision applications," \textit{arXiv preprint arXiv:1704.04861}, 2017.

\bibitem{sandler2018}
M. Sandler, A. Howard, M. Zhu, A. Zhmoginov, and L. C. Chen, "MobileNetV2: Inverted residuals and linear bottlenecks," \textit{Proceedings of the IEEE conference on computer vision and pattern recognition}, pp. 4510-4520, 2018.

\bibitem{everingham2010}
M. Everingham, L. Van Gool, C. K. Williams, J. Winn, and A. Zisserman, "The pascal visual object classes (voc) challenge," \textit{International journal of computer vision}, vol. 88, no. 2, pp. 303-338, 2010.

\bibitem{lin2014}
T. Y. Lin, M. Maire, S. Belongie, J. Hays, P. Perona, D. Ramanan, P. Dollár, and C. L. Zitnick, "Microsoft coco: Common objects in context," \textit{European conference on computer vision}, pp. 740-755, 2014.

\bibitem{geiger2012}
A. Geiger, P. Lenz, and R. Urtasun, "Are we ready for autonomous driving? the kitti vision benchmark suite," \textit{2012 IEEE conference on computer vision and pattern recognition}, pp. 3354-3361, 2012.

\bibitem{caesar2020}
H. Caesar, V. Bankiti, A. H. Lang, S. Vora, V. E. Liong, Q. Xu, A. Krishnan, Y. Pan, G. Baldan, and O. Beijbom, "nuScenes: A multimodal dataset for autonomous driving," \textit{Proceedings of the IEEE/CVF conference on computer vision and pattern recognition}, pp. 11621-11631, 2020.

\bibitem{cordts2016}
M. Cordts, M. Omran, S. Ramos, T. Rehfeld, M. Enzweiler, R. Benenson, U. Franke, S. Roth, and B. Schiele, "The cityscapes dataset for semantic urban scene understanding," \textit{Proceedings of the IEEE conference on computer vision and pattern recognition}, pp. 3213-3223, 2016.

\bibitem{zhou2017}
B. Zhou, A. Lapedriza, A. Khosla, A. Oliva, and A. Torralba, "Places: A 10 million image database for scene recognition," \textit{IEEE transactions on pattern analysis and machine intelligence}, vol. 40, no. 6, pp. 1452-1464, 2017.

\bibitem{deng2009}
J. Deng, W. Dong, R. Socher, L. J. Li, K. Li, and L. Fei-Fei, "ImageNet: A large-scale hierarchical image database," \textit{2009 IEEE conference on computer vision and pattern recognition}, pp. 248-255, 2009.

\bibitem{papandreou2017}
G. Papandreou, I. Kokkinos, and P. A. Savalle, "Modeling local and global deformations in deep learning: epitomic convolution, multiple instance learning, and sliding window detection," \textit{Proceedings of the IEEE conference on computer vision and pattern recognition}, pp. 390-399, 2015.

\bibitem{kingma2014}
D. P. Kingma and J. Ba, "Adam: A method for stochastic optimization," \textit{arXiv preprint arXiv:1412.6980}, 2014.

\bibitem{paszke2019}
A. Paszke, S. Gross, F. Massa, A. Lerer, J. Bradbury, G. Chanan, T. Killeen, Z. Lin, N. Gimelshein, L. Antiga, A. Desmaison, A. Kopf, E. Yang, Z. DeVito, M. Raison, A. Tejani, S. Chilamkurthy, B. Steiner, L. Fang, J. Bai, and S. Chintala, "PyTorch: An imperative style, high-performance deep learning library," \textit{Advances in neural information processing systems}, vol. 32, pp. 8026-8037, 2019.

\bibitem{abadi2016}
M. Abadi, P. Barham, J. Chen, Z. Chen, A. Davis, J. Dean, M. Devin, S. Ghemawat, G. Irving, M. Isard, M. Kudlur, J. Levenberg, R. Monga, S. Moore, D. G. Murray, B. Steiner, P. Tucker, V. Vasudevan, P. Warden, M. Wicke, Y. Yu, and X. Zheng, "TensorFlow: A system for large-scale machine learning," \textit{12th USENIX symposium on operating systems design and implementation}, pp. 265-283, 2016.

\end{thebibliography}
